Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-07-13 prop_letter.tex
Some updates from Gerardo...
Updating by Gerardo...

In this paper, we extend our previous work on the foundations of
computability in quantum physics by exploring the application of
interval-valued probability measures (IVPM) to achieving a coherent
formulation of finite-resource quantum mechanics.  In particular, we quantify the concept of imperfect quantum measurement and its impact on the interpretation of the Kochen-Specker and Gleason theorems and their implications for the foundations of quantum mechanics.

@@

Here we develop a mathematical framework that includes the uncertainties of finite precision and imperfections in the quantum measurement process.  Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems\nb\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable by
extending classical interval-valued probability
measures\nb\cite{JamisonLodwick2004} to the quantum domain.  We begin by axiomatizing a quantum interval-valued probability measure (QIVPM) framework, which will provide a quantitative description of the impact of finite precision quantum measurements.

In particular, the effect of finite precision on contextuality and the
Kochen-Specker theorem is shown to have a resolution in terms of our quantum interval-valued probability measure framework.  To proceed, we use a formulation of the Mermin square that uses the
expectation value of an observable with respect
to a QIVPM, where the interval-valued
expectation value is consistent with the expectation value of the
same observable on continuous quantum mechanics. Furthermore, the
Kochen-Specker theorem is robust in our framework in a sense that
it would remain valid when the experimental error is less than a particular
bound.

A key aspect of quantum measurement is its stochastic nature as reflected
in Born's rule\nb\citep{Born1983,peres1995quantum,544199}. Gleason's
theorem establishes the existence of a unique state (in Hilbert spaces
of dimension $\ge3$), the probabilities of whose outcomes are computed
from Born's rule assuming infinite precision\nb\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Then it is legitimate to question what happens when infinite precision
cannot be achieved. While a QIVPM\nb$\bar{\mu}$ might not be consistent
with a unique state\nb$\rho'$ on all projectors on a Hilbert space,
if $\bar{\mu}$ is induced by a mixed state\nb$\rho$, $\rho'$ would
close to $\rho$ when we shrink the length of intervals under some
good conditions, and recover original Gleason's theorem in the limiting
case.

The paper is organized as follows.   In Sec.\ \ref{sec:fuzzy} we begin by introducing the foundations of quantum probability space and fuzzy measurement.  We then move on to introduce the concept of quantum interval valued probability in Sec.\ \ref{sec:Interval-Uncertainty}.  This QIVPM will provide the mathematical framework to quantify the impact of finite precision measurements on quantum mechanics.  For instance, in Sec.\ \ref{sec:Kochen-Specker},  we quantify the domain of validity of a  contextual measurement, thus addressing the conditions under which the Kochen-Specker theorem applies.  This provides a way of resolving the Meyer-Mermin debate, which is the main result of our Theorem \ref{cor:Kochen-Specker-IVPM}.  In Sec.\ \ref{sec:Gleason}, we study Gleason's theorem in the context of imprecise measurements, concluding that QIVPM provides a framework with quantitative bounds in which Gleason's theorem, while formally invalid, holds asymptotically.  Finally, we present our conclusions in Sec.~\ref{sec:Conclusion}.

>>>>

@@

In this paper, we extend our previous work on the foundations of
computability in quantum physics by exploring the application of
interval-valued probability measures (IVPM) to achieving a coherent
formulation of finite-resource quantum mechanics.  To this end, we develop a mathematical framework that includes the uncertainties of finite precision and imperfections in the quantum measurement process. In particular, we quantify the concept of imperfect quantum measurement and its impact on the interpretation of the Kochen-Specker and Gleason theorems and their implications for the foundations of quantum mechanics.  Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems\nb\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable by
extending classical interval-valued probability
measures\nb\cite{JamisonLodwick2004} to the quantum domain.  We thus begin by axiomatizing a quantum interval-valued probability measure (QIVPM) framework.

The effect of finite precision and imperfections on contextuality and the
Kochen-Specker theorem is shown to have a resolution in terms of our QIVPM framework.  
To proceed, we use a formulation of the ``Mermin-Peres square'' \nb\cite{peres1995quantum,Mermin1990Simple} that uses the
expectation value of a physical observable with respect
to a QIVPM, where the interval-valued
expectation value is consistent with the expectation value of the
same observable in conventional quantum mechanics with infinite resources. We show that, 
as long as errors and uncertainties remain smaller that a particular value, the
Kochen-Specker theorem remains valid. In other words, we established a rigorous bound for 
its validity and, therefore, the limits for physical reality to remain contextual.

A key aspect of a quantum measurement is its statistical nature, as reflected
in the application of the Born rule to determine the probability of a given measurement 
outcome \nb\citep{Born1983,peres1995quantum,544199}. The Gleason
theorem establishes that there is no alternative to the Born rule by 
demonstrating, using reasonable continuity arguments in Hilbert spaces
of dimension $d\ge3$, the existence of a unique state 
$\rho$ consistent with the statistical predictions computed from 
that rule \nb\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
It is then legitimate to question what happens when infinite precision
cannot be achieved. We will show that, 
while a QIVPM  might not be consistent
with a unique state\nb$\rho$ on all projectors defined on a Hilbert space of dimension $d\ge3$,
if the QIVPM is induced by some mixed state\nb$\rho'$ ,

$\rho'$ would
close to $\rho$ when we shrink the length of intervals under some
good conditions, and recover the original Gleason theorem in the limiting
case.

The paper is organized as follows.   In Sec.\ \ref{sec:fuzzy} we begin by introducing the foundations of quantum probability space and fuzzy measurement.  We then move on to introduce the concept of quantum interval valued probability in Sec.\ \ref{sec:Interval-Uncertainty}.  This QIVPM will provide the mathematical framework to quantify the impact of finite precision measurements on quantum mechanics.  For instance, in Sec.\ \ref{sec:Kochen-Specker},  we quantify the domain of validity of a  contextual measurement, thus addressing the conditions under which the Kochen-Specker theorem applies.  This provides a way of resolving the Meyer-Mermin debate, which is the main result of our Theorem \ref{cor:Kochen-Specker-IVPM}.  In Sec.\ \ref{sec:Gleason}, we study Gleason's theorem in the context of imprecise measurements, concluding that QIVPM provides a framework with quantitative bounds in which Gleason's theorem, while formally invalid in a universe with bounded resources, holds asymptotically in a mathematically precise way.  Finally, we present our conclusions in Sec.~\ref{sec:Conclusion}.
